Modeling

### 2020-10-19 police_eda.Rmd
eda

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tidyverse)
```

@@

all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
  mutate(b = as.numeric(b > 0),
         w = as.numeric(w > 0),
         l = as.numeric(l > 0),
         a = as.numeric(a > 0),
         so =  as.numeric(str_count(all$SubjectRace, "O") > 0),
         smale = as.numeric(str_count(all$SubjectGender, "M") > 0),
         sfem = as.numeric(str_count(all$SubjectGender, "F") > 0),
         ob = as.numeric(str_count(all$OfficerRace, "B") > 0),
         ow = as.numeric(str_count(all$OfficerRace, "W") > 0),
         ol = as.numeric(str_count(all$OfficerRace, "L") > 0),
         oa = as.numeric(str_count(all$OfficerRace, "A") > 0),
         oo =  as.numeric(str_count(all$OfficerRace, "O") > 0),
         omale = as.numeric(str_count(all$OfficerGender, "M") > 0),
         ofem = as.numeric(str_count(all$OfficerGender, "F") > 0),
         other = as.numeric(other), 
         replica = as.numeric(replica), 
         knife = as.numeric(knife), 
         gun = as.numeric(weapon == "gun"), 
         unarmed = as.numeric(weapon == "unarmed"),
         unknown_weapon = as.numeric(weapon == "U"),
         fatal = as.numeric(str_count(all$Fatal, "F") > 0)) %>% 
  select(-c(weapon, SubjectGender, OfficerGender))

@@

race_count <- ggplot(temp, aes(x = type, y = count, fill = fatal)) +
  geom_bar(position = "dodge", stat="identity") + 
  labs(x = "Present Victim Race", y = "Count", title = "Counts for Fatal vs. Non-Fatal Single-Victim Police Shootings Across Different Races") +
  theme(legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

race_prop <- ggplot(temp, aes(x = type, y = prop, fill = fatal)) +
  geom_bar(position = "stack", stat="identity") + 
  labs(x = "Present Victim Race", y = "Proportion", title = "Proportions for Fatal vs. Non-Fatal Single-Victim Police Shootings Across Different Races") +
  theme(legend.text = element_text(size = 7),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

@@

race_count1 <- ggplot(temp, aes(x = type, y = count, fill = fatal)) +
  geom_bar(position = "dodge", stat="identity") + 
  labs(x = "Present Victim Race", y = "Count", title = "Counts for Fatal vs. Non-Fatal Multiple-Victim Police Shootings Across Different Races") +
  theme(legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

race_prop1 <- ggplot(temp, aes(x = type, y = prop, fill = fatal)) +
  geom_bar(position = "stack", stat="identity") + 
  labs(x = "Present Victim Race", y = "Proportion", title = "Proportions for Fatal vs. Non-Fatal Multiple-Victim Police Shootings Across Different Races") +
  theme(legend.text = element_text(size = 7),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

@@

```{r fatally shot prop for all shootings of varying numbers of victims}
temp <- final_eda %>% 
  group_by(NumberOfSubjects) %>% 
  summarise(prop = sum(fatal == "Fatal")/n(),
            total = n())
temp
```

```{r black to white ratio}
temp <- final_eda %>% 
  group_by(city, fatal) %>% 
  summarise(btw = black/white) %>% 
  unique

plot <- ggplot(temp, aes(x = btw, y = fatal, fill = fatal)) + 
    geom_boxplot()  + 
  labs(x = "Black to White Ratio", y = "Fatal vs. Non-Fatal Shooting", title = "City Black to White Ratio Distributions for Fatal vs. Non-Fatal Police Shootings") +
  theme(legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5))

ggsave(filename = "figures/btweda.png", plot)
```
Means are not that different, will probably just use percent black and percent white then.

>>>>

@@

all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
  mutate(b = as.numeric(b > 0),
         w = as.numeric(w > 0),
         l = as.numeric(l > 0),
         a = as.numeric(a > 0),
         so =  as.numeric(str_count(all$SubjectRace, "O") > 0),
         smale = as.numeric(str_count(all$SubjectGender, "M") > 0),
         sfem = as.numeric(str_count(all$SubjectGender, "F") > 0),
         ob = as.numeric(str_count(all$OfficerRace, "B") > 0),
         ow = as.numeric(str_count(all$OfficerRace, "W") > 0),
         ol = as.numeric(str_count(all$OfficerRace, "L") > 0),
         oa = as.numeric(str_count(all$OfficerRace, "A") > 0),
         oo =  as.numeric(str_count(all$OfficerRace, "O") > 0),
         omale = as.numeric(str_count(all$OfficerGender, "M") > 0),
         ofem = as.numeric(str_count(all$OfficerGender, "F") > 0),
         other = as.numeric(other), 
         replica = as.numeric(replica), 
         knife = as.numeric(knife), 
         gun = as.numeric(weapon == "gun"), 
         unarmed = as.numeric(weapon == "unarmed"),
         unknown_weapon = as.numeric(weapon == "U"),
         fatal = as.numeric(str_count(all$Fatal, "F") > 0)) %>% 
  select(-c(weapon))

@@

race_count <- ggplot(temp, aes(x = type, y = count, fill = fatal)) +
  geom_bar(position = "dodge", stat="identity") + 
  labs(x = "Victim Race", y = "Count", title = "Counts for Fatal vs. Non-Fatal Single-Victim Police Shootings Across Different Races") +
  theme(legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

race_prop <- ggplot(temp, aes(x = type, y = prop, fill = fatal)) +
  geom_bar(position = "stack", stat="identity") + 
  labs(x = "Victim Race", y = "Proportion", title = "Proportions for Fatal vs. Non-Fatal Single-Victim Police Shootings Across Different Races") +
  theme(legend.text = element_text(size = 7),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

@@

race_count1 <- ggplot(temp, aes(x = type, y = count, fill = fatal)) +
  geom_bar(position = "dodge", stat="identity") + 
  labs(x = "Victim Race", y = "Count", title = "Counts for Fatal vs. Non-Fatal Multiple-Victim Police Shootings Across Different Races") +
  theme(legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

race_prop1 <- ggplot(temp, aes(x = type, y = prop, fill = fatal)) +
  geom_bar(position = "stack", stat="identity") + 
  labs(x = "Victim Race", y = "Proportion", title = "Proportions for Fatal vs. Non-Fatal Multiple-Victim Police Shootings Across Different Races") +
  theme(legend.text = element_text(size = 7),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

@@

```{r fatally shot prop for all shootings of varying numbers of victims}
temp <- final_eda %>% 
  group_by(NumberOfSubjects) %>% 
  summarise(prop = sum(fatal == "Fatal")/n(),
            total = n())
temp
```
For sample size > 10, fatality seems to be lower for multiple victims rather than just 1.

```{r black to white ratio}
temp <- final_df %>% 
  group_by(city) %>% 
  summarise(btw = black/white, propfatal = sum(fatal)/n()) %>% 
  unique

plot <- ggplot(temp, aes(x = btw, y = propfatal)) + 
    geom_point()  + 
  labs(x = "Black to White Ratio", y = "Fatal Shootings Proportions", title = "Proportion of Fatal Shootings vs. Black to White Ratios in Cities") +
  theme(legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5))

ggsave(filename = "figures/btweda.png", plot)
```

Can't really tell if there's a distribution here, maybe negatively correlated.

```{r black percent}
temp <- final_df %>% 
  group_by(city) %>% 
  summarise(bprop = black/city_pop_2013, propfatal = sum(fatal)/n()) %>% 
  unique

plot <- ggplot(temp, aes(x = bprop, y = propfatal)) + 
    geom_point()  + 
  labs(x = "Black Percentage of Population", y = "Fatal Shootings Proportions", title = "Proportion of Fatal Shootings vs. Black Population Percentages in Cities") +
  theme(legend.text = element_text(size = 8),
        legend.key.size = unit(0.4, "cm"),
        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
        plot.title = element_text(size=10, hjust = 0.5))

ggsave(filename = "figures/bpropeda.png", plot)
```

Same thing, can't really tell if there's a distribution here, maybe negatively correlated.

```{r}
temp <- final_df %>% 
  summarise(Black = sum(str_count(OfficerRace, "B")), 
            White = sum(str_count(OfficerRace, "W")),
            Hispanic = sum(str_count(OfficerRace, "L")),
            Asian = sum(str_count(OfficerRace, "A")), 
            Other = sum(str_count(OfficerRace, "O")))
ggplot(temp, aes(x = ))
```

```{r}
temp <- final_df %>% 
  filter(sb == 1) %>% 
  group_by(fatal) %>% 
  summarise(ob = sum(ob == 1),
            ow = sum(ow == 1),
            ol = sum(ol == 1),
            oa = sum(oa == 1),
            oo = sum(oo == 1)) %>% 
  mutate(ob = ob/(ob + ow + ol + oa + oo), 
         ow = ow/(ob + ow + ol + oa + oo),
         ol = ol/(ol + ow + ol + oa + oo), 
         oa = oa/(ob + ow + ol + oa + oo), 
         oo = oo/(ob + ow + ol + oa + oo))
temp <- rename(temp, Black = ob, White = ow, Latino = ol, Asian = oa, Other = oo)

temp2 <- temp %>%
    t() %>%
    as.data.frame(stringsAsFactors = F) %>%
    rownames_to_column("race") %>%
    `rownames<-`(NULL)
temp2 <- rename(temp2, "Non-Fatal" = 2, Fatal = 3)
temp2 <- temp2[-1,]
temp2 <- melt(temp2)

officerplot <- ggplot(temp2, aes(x=race, y = value, fill = variable)) +
  geom_bar(position = "dodge", stat="identity") + 
  labs(x = "Officer Race", y = "Proportion", title = "Race Proportions for Fatal vs. Non-Fatal Police Shootings Where Black Victims were Present") +
  theme(legend.text = element_text(size = 7),
        legend.key.size = unit(0.4, "cm"),
        plot.title = element_text(size=10, hjust = 0.5)) + 
  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

ggsave(filename = "figures/officerpropsb.png", officerplot)
```

### 2020-10-19 police_modeling.Rmd
eda

## Modeling
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tidyverse)
```

```{r load-data}
all <- read.csv("allshootingsdata.csv")

philly <- read.csv("philly_shootings.csv")
cincinnati <- read.csv("cincinnati_shootings.csv")
```

```{r clean-all}
all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "b", "l", "w", "a", "tr", "AvgAge", "VA", "F", "nshots", "sit"))

all <- all %>% 
  separate_rows(SubjectRace, SubjectGender, Fatal, sep = ";")

officer_gender <- sapply(str_split(all$OfficerGender, ";"), length)
officer_race <- sapply(str_split(all$OfficerRace, ";"), length)
out <- which((officer_gender == officer_race) == FALSE)
all <- all[-out,]

all <- all %>% 
  separate_rows(OfficerRace, OfficerGender, sep = ";") %>% 
  separate_rows(OfficerRace, OfficerGender, sep = ",")

officer_gender2 <- sapply(str_split(all$OfficerGender, "/"), length)
officer_race2 <- sapply(str_split(all$OfficerRace, "/"), length)
out2 <- which((officer_gender2 == officer_race2) == FALSE)
all <- all[-out2,]

all$OfficerRace <- trimws(all$OfficerRace)
all$OfficerGender <- trimws(all$OfficerGender)
all$Fatal <- trimws(all$Fatal)

all <- all %>% 
  filter(SubjectRace != "U") %>% 
  filter(OfficerRace != "U") %>% 
  filter(OfficerRace != "Unknown") %>% 
  filter(OfficerRace != "NA") %>% 
  filter(OfficerRace != "") %>% 
  filter(OfficerRace != "I") %>% 
  filter(OfficerGender != "U") %>% 
  filter(OfficerGender != "N") %>% 
  filter(SubjectGender != "U") %>% 
  filter(SubjectGender != "N") %>% 
  filter(Fatal != "U")

all$OfficerRace[all$OfficerRace == "WHITE"] <- "W"
all$OfficerRace[all$OfficerRace == "BLACK"] <- "B"
all$OfficerRace[all$OfficerRace == "ASIAN"] <- "A"

all$OfficerGender[all$OfficerGender == "FEMALE"] <- "F"
all$OfficerGender[all$OfficerGender == "MALE"] <- "M"
all$OfficerGender[all$OfficerGender == "Male"] <- "M"
```

>>>>

## Modeling
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tidyverse)
library(lmtest)
```

```{r load-data}
final_df <- readRDS("data/final.rds")

single_victims <- final_df %>% 
  filter(NumberOfSubjects <= 1)

mult_victims <- final_df %>% 
  filter(NumberOfSubjects > 1)
```

```{r}
fit1 <- glm(formula = fatal ~ SubjectRace, family = binomial, data=single_victims)

fit2 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013), family = binomial, data=single_victims)
lrtest(fit2, fit1)
summary(fit2)

# Victim Gender doesn't appear to be significant
fit3 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + SubjectGender, family = binomial, data=single_victims)
lrtest(fit3, fit2)

fit4 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob, family = binomial, data=single_victims)
lrtest(fit3, fit4)

fit5 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow, family = binomial, data=single_victims)
lrtest(fit5, fit4)

fit6 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol, family = binomial, data=single_victims)
lrtest(fit5, fit6)

fit7 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace, family = binomial, data=single_victims)
lrtest(fit6, fit7)

fit8 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace, family = binomial, data=single_victims)
lrtest(fit8, fit7)

fit9 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace + SubjectArmed, family = binomial, data=single_victims)
lrtest(fit8, fit9)
```
